---
title: "lucerne_quality_of_life_clustering"
author: 'Hackathon HSLU'
date: "28 November 2020"
output:
  html_document: 
    toc: TRUE
    toc_float: TRUE
  pdf_document:
    toc: TRUE
---

# Setup

```{r setup, include=FALSE}
library("knitr")
knitr::opts_chunk$set(echo = TRUE)
library("stats")
library("dplyr")
library("ggplot2")
library("ggfortify")
library("tidyverse")

df <- read_csv2("cleaned_data.csv")
colnames(df)
set.seed(1)

```

Import data and look at columns.

```{r import, include=FALSE}
df <- read_csv2("cleaned_data.csv")
colnames(df)

```

# Preprocessing

The data is prepared for analysis.

```{r preprocessing, echo=TRUE}
#drop
drop2 <- c("lfdn", "v_238", "v_211", "quality", "v_360", "v_205" , "v_290", "v_88", "v_363")
df2 <- dplyr::select(df, -drop2)

#remove rows where survey was not completed
df2 <- subset(df2, df2$v_321 != -77)

#substitute value
df2$v_83[df2$v_83 == -77] <- 0
df2$v_84[df2$v_84 == -77] <- 0
df2$v_85[df2$v_85 == -77] <- 0
df2$v_86[df2$v_86 == -77] <- 0
df2$v_87[df2$v_87 == -77] <- 0
df2$v_104[df2$v_104 == -77] <- 0
df2$v_105[df2$v_105 == -77] <- 0

#drop columns where not enough answers provided
drop3 <- c("v_73", "v_74", "v_75", "v_152", "v_153", "v_154", "v_269", "v_274", "v_275", "ArbeitsplatzLuzern", "v_89", "v_90")
df2 <- dplyr::select(df2, -drop3)

#drop columns because doesnt work in PCA
drop4 <- c("lastpage", "duration", "X1")
df2 <- dplyr::select(df2, -drop4)
```

# Analysis with clusters

## Principal component analysis

```{r PCA, echo=TRUE}
library("MASS")
pca <- princomp(df2, scores=TRUE)
plot(pca,main="scree plot")
summary(pca)
```

## Clustering

```{r cluster, echo=TRUE}
## Scale data
df2_scaled <- scale(df2, center = FALSE, scale = TRUE)
boxplot(df2_scaled)

## Distance matrix (used below)
dp <- dist(df2_scaled)

## Choose number of clusters k with scree plot
wss <- rep(0, 20)
for (i in 1:20) wss[i] <- sum(kmeans(df2, centers = i, nstart = 20)$withinss)
par(mfrow = c(1,1))
plot(1:20, wss, type = "b", xlab = "Number of groups", ylab = "Within groups sum of squares") ## 3 centers was a good choice
## Note: If nstart=1 is used, the results can vary because of random starting configurations in kmeans

## k-means 
ckm <- kmeans(df2, centers = 5, nstart = 10)
grpsKM <- ckm$cluster
```

There is no clear elbow point, we choose 5 groups. 

```{r, echo=TRUE}
## k-means 
set.seed(1)
ckm <- kmeans(df2, centers = 5, nstart = 10)
grpsKM <- ckm$cluster
```

## View Clusters on Pricipal components

```{r cluster-view, echo=TRUE}
pc <- princomp(df2)
par(mfrow = c(1,1))

# visualize in PC 1 & 2
pr <- pc$scores[,1:2]
plot(pr, pch = grpsKM, col=grpsKM, lwd=2)
legend("bottomright", legend = 1:5, pch = 1:5, col=1:5, bty="n")
title(main = "PC1 vs PC2")

# visualize in PC  2 & 3
pr <- pc$scores[,2:3]
plot(pr, pch = grpsKM, col=grpsKM, lwd=2)
legend("bottomright", legend = 1:5, pch = 1:5, col=1:5, bty="n")
title(main = "PC2 vs PC3")

# visualize in PC  1 & 3
pr <- cbind("Comp.1" = pc$scores[,1], "Comp.3" = pc$scores[,3])
plot(pr, pch = grpsKM, col=grpsKM, lwd=2)
legend("bottomright", legend = 1:5, pch = 1:5, col=1:5, bty="n")
title(main = "PC1 vs PC3")

```


# Select variables and build personas

## Mean of data by cluster

```{r, echo=TRUE}
df.means <- aggregate(df2, by=list(cluster=grpsKM), mean)
```


## Analysis of variance

Here we analyse the relevant variables, meaning the ones that differ significantly between the clusters.

```{r, echo=TRUE}
df2_scaled <- as.data.frame(df2_scaled)
df3 <- df2_scaled
df3["cluster"] <- grpsKM


fit <- aov(cluster ~ .-cluster, data=df3)
summary(fit)
```

## Select and rename significant variables

```{r, echo=TRUE}

#note: this selection was based on a previous clustering, it will not match with the current analysis


selection.aov <- c("Lebensqualallg",
                   "Sex",
                   "Altergruppe",
                   "Quartier",
                   "Zuzug", 
                   "Erwerb",
                   "v_364",
                   "v_79",
                   "v_357",
                   "v_359",
                   "v_86",
                   "v_56",
                   "v_305",
                   "v_306",
                   "v_41",
                   "v_42",
                   "v_50",
                   "v_51",
                   "v_187",
                   "v_279",
                   "v_280",
                   "v_324",
                   "v_325",
                   "v_316",
                   "v_106",
                   "v_212",
                   "v_366",
                   "v_248")


selection.aov2 <- c("Lebensqualallg",
                   "Sex",
                   "Altergruppe",
                   "Quartier",
                   "Zuzug", 
                   "Erwerb",
                   "v_294",
                   "v_79",
                   "v_355",
                   "v_357",
                   "v_302",
                   "v_311",
                   "v_39",
                   "v_323",
                   "v_325",
                   "v_106",
                   "v_212",
                   "v_366",
                   "v_367",
                   "v_293",
                   "v_256")


df.aov.selected <- dplyr::select(df.means, selection.aov2)

df.aov.selected.transposed <- as.data.frame(t(df.aov.selected))
print(df.aov.selected.transposed)
```

## Plot Kinder pro Cluster

```{r, echo=TRUE}
df2["cluster"] <- grpsKM
```

```{r, echo=TRUE}
ggplot(df2, aes(x=df2$cluster, fill = as.factor(df2$Kinder_Schule)))+
  geom_bar()+ 
  labs(fill='Kinder')+
  scale_fill_discrete(labels = c("schulpflichtige Kinder", "keine schulpflichtige Kinder"))
```

## Plot Haushalt pro Cluster

```{r, echo=TRUE}
ggplot(df2, aes(x=df2$cluster, fill = as.factor(df2$HH)))+
  geom_bar()+ 
  labs(fill='Haushalt')+
  scale_fill_discrete(labels = c("Einpersonenhaushalt", "Elternteil mit Kinder", "Paar ohne Kinder", "Paar mit Kinder", "anderer Mehrpersonenhaushalt"))
```

## Plot Alter pro Cluster

```{r, echo=TRUE}
ggplot(df2, aes(x=df2$cluster, fill = as.factor(df2$Altergruppe)))+
  geom_bar()+ 
  labs(fill='Altersgruppe')+
  scale_fill_discrete(labels = c("0-9", "10-19", "20-29", "30-39", "40-49", "50-65", "65-84", "85+"))
```


## Export

```{r, echo=TRUE}
write.csv2(df.aov.selected, "df_aov_selected2.csv")
write.csv2(df.aov.selected.transposed, "df_aov_selected_transposed2.csv")
```


